Unifying Gompertzian growth with the communicable disease spreading paradigm

### 2022-04-27 main.tex
mod last par on sec 2 + equations

Thus, a more direct interpretation of the observations is that mortality was caused by a planetary perturbation, modeled as a random process, to which organisms gradually develop resistance at a geometric rate \citep{boxenbaum2017hypotheses,neafsey1988gompertz}. 
Here, a geometric rate is an exponential rate in the log-transformed domain, which is the natural transformation for many processes in nature \citep{zhang1994log}. 
The log-normal distribution of the abundance of $F(t)$ can be obtained by directly solving the stochastic equation in \ref{eq:microscopicOriginal} \citep{skiadas2010exact,petroni2020gompertz}, or from thermodynamic principles \citep{sitaram1984statistical,gunasekaran1982lon}. 
Intuitively, this is seen by noting that the solution to the perturbation model in the log-domain (\ref{eq:microscopic}) is Gaussian in the variable $Z(t)$, thus suggesting a log-normal distribution of $F(t)$.

@@

\begin{align}
\label{eq:compare}
\begin{split}
\dot{F} = \frac{d}{dt}F(t) = \beta F(t) (1-F(t)) \quad&\quad \text{Logistic}\\ 
\frac{d}{dt}\ln{(F(t))} = -\beta \ln{(F(t)}) &\quad \text{Gompertz},
\end{split}
\end{align}
where $F$ is here deterministic.

@@

A microscopic solution could be modeled by splitting the system into $N$ microscopic deterministic units, $F(t) \rightarrow f_1(t), f_2(t),$ etc, from which macroscopic growth would be obtained by taking the average over the individual units,
\begin{equation}
F(t) = \frac{1}{N}\sum_i^N f_i(t),
\end{equation}
where the lower case $f_i$ emphasize the microscopic quality of the variables, and is here interpreted as probability of infection.

One could add network interaction necessitating a corresponding matrix version of (\ref{eq:compare})
\begin{equation}
\label{eq:networkSIR}
\frac{d f_i}{dt} = \beta (1-f_i)\sum_j{a_{ij}}f_j \quad \forall i,
\end{equation}
using the shorthanded $f_i=f_i(t)$ and with a fixed correlation governed by the network's growth rate, $b$, and adjacency matrix, $\{a_{i,j}\}=\mathbf{A} \in \mathbb{R}^{N \times N}$, a binary matrix with ones where the $i^{th}$ and $j^{th}$ nodes are connected, and zeroes otherwise. Notice here that there is an implied causality going from the infected to susceptible, which will become relevant below. Furthermore, a linear correlation between variables is seen as the partial derivatives of the instantaneous growth rate with respect to pairs of microscopic variables,
\begin{equation}
\frac{\partial^2 \dot{F}}{\partial f_i \partial f_j} = -\frac{\beta}{N} a_{i,j}.
\end{equation}
Still, no Gompertz curve will emerge at the onset of the growth process. \citet{estrada2022networked} provide illuminating analysis on this topic.

>>>>

Thus, a more parsimonious interpretation of the observations not reliant on a transmission network is that mortality was caused by a planetary perturbation, modeled as a random process, to which organisms gradually develop resistance at a geometric rate in the log-transformed domain \citep{boxenbaum2017hypotheses,neafsey1988gompertz}, which is the natural transformation for many processes in nature \citep{zhang1994log}. 
Under this model, the distribution of the abundance of $F(t)$ is log-normal, a result that can be obtained by directly solving the stochastic equation in \ref{eq:microscopicOriginal} \citep{skiadas2010exact,petroni2020gompertz}, or from thermodynamic principles \citep{sitaram1984statistical,gunasekaran1982lon}. 
Intuitively, this is seen by noting that the solution to the perturbation model in the log-domain (\ref{eq:microscopic}) is Gaussian in the variable $Z(t)$, thus suggesting a log-normal distribution of $F(t)$.

@@

\begin{subequations}
\begin{align}
\dot{F} = \frac{d}{dt}F(t) & = \beta F(t) (1-F(t)) \quad \text{Logistic}\label{eq:compareLog}\\ 
\frac{d}{dt}\ln{(F(t))} & = -\beta \ln{(F(t)}) \quad\quad\text{Gompertz}\label{eq:compareGom},
\end{align}
\end{subequations}
where $F$ is here deterministic.

@@

A microscopic solution could be modeled by splitting the system into $N$ microscopic deterministic units, $F(t) \rightarrow f_1(t), f_2(t), ..., f_N(t)$, where the lower case $f_i$ emphasize the microscopic quality of the variables, and is here interpreted as probability of infection. From these microscopic units, macroscopic growth would be obtained by taking their average,
\begin{equation}
F(t) = \frac{1}{N}\sum_i^N f_i(t),
\end{equation}

One could add network interaction necessitating a corresponding matrix version of (\ref{eq:compareLog})
\begin{equation}
\label{eq:networkSIR}
\frac{d f_i}{dt} = \beta (1-f_i)\sum_j{a_{ij}}f_j \quad \forall i,
\end{equation}
using the shorthanded $f_i=f_i(t)$ and with a fixed correlation governed by the network's growth rate, $b$, and adjacency matrix, $\{a_{i,j}\}=\mathbf{A} \in \mathbb{R}^{N \times N}$, a binary matrix with ones where the $i^{th}$ and $j^{th}$ nodes are connected, and zeroes otherwise. Notice here that there is an implied causality going from the infected to susceptible, which will become relevant below. Furthermore, a linear correlation between variables is seen as the partial derivatives of the instantaneous growth rate with respect to pairs of microscopic variables,
\begin{equation}
\frac{\partial^2 \dot{F}}{\partial f_i \partial f_j} = -\frac{\beta}{N} a_{i,j}.
\end{equation}
Still, no Gompertz curve will emerge at the onset of the growth process. \citet{estrada2022networked} provide illuminating analysis on this topic.
